Modeling HEPES buffer with a charge-balance approach

### 2023-11-15 manuscript.Rmd
Added reference to Nash/Bhattacharjee and an Appendix about methods.

>>>>

Appendix: Software considerations
=================================

The `nlsr` package @CRAN:nlsr which has been used for the estimation of
parameters in the nonlinear models for this article has some features that 
more easily permitted the computations. Moreover, discussions concerning how
to use the software led to unexpected possibilities in that use.

### Ensuring parameters follow a specific ordering

We require that the three $\mathrm{p}K_a$ values be in increasing order of
numerical value. To simplify the discussion here, we will write these as 
`p1`, `p2` and `p3` within our R program scripts. Thus we want

```
       p1  <  p2  < p3
```

One approach, which we have used here, is to reparameterize the formulas for
the model so

``` 
      p2  =  p1 + d12
      p3  =  p1 + d12 + d23
```

If we constrain `d12` and `d23` to be positive, or at least non-negative, then
the new parameterization will ensure the ordering. However, this does require 
that the software can offer bounds constraints on the parameters. `nslr` is 
capable in this way, and even allows upper and lower bounds to be equal, 
resulting in fixed parameters, or **masks**. Other software in R has similar
capabilities, but the built-in `nls()` function only handles bounds if the
`port` algorithm is used, and there are warnings that this code may be 
incomplete. Package `minpack.lm` also handles bounds, though a failure was
encountered in one example (@NashBhattacharjee).

### Vectorizing a function ???

JN: As yet I don't understand the workings of this, and must admit to never
having used Vectorize(). I think it worthwhile giving a good explanation.

### Gauging the reliability of the estimated parameters

All parameter estimation methods must contend with the possibility that 
a variety of parameter values may give very similar magnitudes for the loss
function (e.g, sum of squared residuals) that is optimized to obtain the
estimates. In the case of estimation that can be converted to minimizing a
sum of squares, the Jacobian of the vector of residuals will become close 
to singular if parameters are indeterminate. This can be judged by computing
the **singular values** (??JN: do we want a reference) of the Jacobian and
package `nlsr` does this and displays the values when a summary of a solution 
is requested. A rule of thumb is to regard the Jacobian as essentially 
singular if the ratio of the smallest to largest singular value is "small".
A value of smaller than 1e-4 would be a typical criterion. Note that more than
one singular value may be small, indicating that there is more than one 
dimension of indeterminacy. However, such indicators apply to the ensemble of
parameters, and as such concern the overall model.
